---
title: "Homework 3"
output: github_document
---

## October 3rd, 2017

### First, always have to load the relevant libraries!

```{r}
library(tidyverse)
library(gapminder)
```

### Maximum and minimum GDP per capita for all continents

```{r}
a <- gapminder %>%
  group_by(continent) %>%
  summarize(MaxGDP=max(gdpPercap), MinGDP=round(min(gdpPercap), 2), DiffGDP=max(gdpPercap)-min(gdpPercap))
knitr::kable(a, format = "markdown")
```

```{r}
a %>%
  ggplot(aes(MaxGDP, MinGDP)) +
  geom_point(aes(color=continent))
```

We can see here that Asia has the biggest discrepancy between a low minimum GDP and a high maximum GDP (also shown in the table!), Africa has the lowest difference because they have the lowest minimum *and* maximum GDP and Oceania also has a small discrepancy in GDP, where they have a much higher minimum GDP than other areas.

### Spread of GDP per capita within the continents

```{r}
b <- gapminder %>%
  group_by(continent) %>%
  summarize(GDP_IQR=IQR(gdpPercap),GDP_SD=sd(gdpPercap), GDP_median_dev=mad(gdpPercap))
knitr::kable(b, format = "markdown")
```

```{r}
ggplot(gapminder, aes(continent, gdpPercap)) +
  geom_boxplot(aes(color=continent))
```

<style type="text/css">
.twoC {width: 100%}
.clearer {clear: both}
.twoC .table {max-width: 50%; float: right}
.twoC img {max-width: 50%; float: left}
</style>

A boxplot showing the spread described by the above table.

```{r}
 ggplot(gapminder, aes(continent, gdpPercap)) +
  geom_boxplot(outlier.shape = NA, aes(color=continent))+
  scale_y_continuous(limits=quantile(gapminder$gdpPercap, c(0.1, 0.991)))
```

The same plot with extreme outliers removed to see the detail here.

### Countries with low life expectancy over time by continent

```{r}
c <- gapminder %>%
  group_by(continent, year) %>%
  filter(lifeExp < 40) %>%
  summarize(numcountries=n_distinct(country))
knitr::kable(c, format="markdown")
```

```{r}
ggplot(c, aes(year, numcountries)) +
  geom_bar(aes(fill=continent), position="dodge", stat="identity")
```

