---
title: "hw02"
output: github_document
---

# Homework Assignment 2

## September 26, 2017

### Exploring Gapminder!

#### First load gapminder and tidyverse to access dplyr
```{r}
library(gapminder)
library(tidyverse)
```

#### "Smell testing" the data
```{r}
is.data.frame(gapminder)
is.matrix(gapminder)
is.vector(gapminder)
is.list(gapminder)
```
Gapminder is a dataframe and also a list but not a matrix or vector

```{r}
class(gapminder)
```
Gapminder's class is a dataframe but also a "tibble"

```{r}
ncol(gapminder)
```
There are 6 columns/variables in the gapminder dataframe

```{r}
nrow(gapminder)
```
There are 1,704 observations/rows within gapminder

```{r}
str(gapminder)
```
The str() function can give the above information in one command!
It also displays the data type of each variable where:
- country and continent are factors
- year and population are integers
- life expectancy and gdp per capita are numbers

```{r}
as_tibble(gapminder) #columns, rows, data types
head(gapminder) #columns, data types
dim(gapminder) #columns and rows
names(gapminder) #columns
length(gapminder) #columns
summary(gapminder) #columns
```
Some alternatives to find number of columns or number of rows
As well as the different data types of the variables
Summary could be very useful if you need max, mins and/or means

#### Exploring individual variables



#### Exploring various plot types



#### Using filter(), select() and %>% 

