---
title: "Powers Vignette"
author: "Vanessa Fladmark"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Powers Vignette"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Powers

The goal of powers is to calculate the exponential power of a number! There are a variety of functions to choose from, such as `square()`, `cube()`, `reciprocal()` and `pow()` for calculating to the power of two, the power of three, the inverse of a power $1/X^a$ and the power of any given number `a`, respectively.

If you already have devtools installed, you can download my package here:

```{r}
devtools::install_github("vanflad/STAT547-hw-fladmark-vanessa/tree/master/Homework%209")
```

## Example

Here are some basic examples which shows you how to solve common power calculations:

```{r}
pow(-2, 4)
square(1:10)
cube(-5)
reciprocal(-2, 3)
```

## Figures

The figures show numbers 1 through 10 and what some common exponential power transformations of that vector look like when displayed visually.

```{r, fig.show='hold', fig.cap = "Original Numbers versus Transformations"}
plot(1:10, square(1:10))
plot(1:10, cube(1:10))
plot(1:10, pow(1:10, 4))
plot(1:10, reciprocal(1:10, 2))
```
