---
title: "hw02"
output: github_document
---

# Homework Assignment 2

## September 26, 2017

### Exploring Gapminder!

#### Bringing rectangular data in

```{r}
library(gapminder)
library(tidyverse)
```

#### "Smell testing" the data

```{r}
is.data.frame(gapminder)
is.matrix(gapminder)
is.vector(gapminder)
is.list(gapminder)
```

Gapminder is a dataframe and a list but not a matrix or vector

```{r}
class(gapminder)
```

Gapminder's class is a dataframe but also a "tibble"

```{r}
ncol(gapminder)
```

There are 6 columns/variables in the gapminder dataframe

```{r}
nrow(gapminder)
```

There are 1,704 observations/rows within gapminder

```{r}
str(gapminder)
```

The str( ) function can give the above information in one command!
It also displays the data type of each variable where:
country and continent are factors, 
year and population are integers, 
and life expectancy and gdp per capita are numbers

```{r}
as_tibble(gapminder) #columns, rows, data types
head(gapminder) #columns, data types
dim(gapminder) #columns and rows
names(gapminder) #columns
length(gapminder) #columns
summary(gapminder) #columns
```

Some alternatives to find number of columns or number of rows, as well as the different data types of the variables, summary could be very useful if you need max, mins and/or means

#### Exploring individual variables

##### Starting with a categorical variable (continent)

```{r}
levels(gapminder$continent)
```

There are five possible values for the different continents

```{r}
summary(gapminder$continent)
table(gapminder$continent) #another command, exact same result
```

Summary( ) shows how frequently each variable appears in dataset
Where Africa has the most observations and Oceania the least

```{r}
barplot(summary(gapminder$continent), cex.names=0.8)
```

A visualization of the same information the summary table gave

##### Then a quantitative variable (life expectancy)

```{r}
summary(gapminder$lifeExp)
```

For numerical data, summary( ) shows min, median, mean, max, etc.

```{r}
hist(gapminder$lifeExp)
```

Visual representation of numerical data is a frequency histogram, displays same info as summary but is easier to see distribution

#### Exploring various plot types

##### Scatterplot (two quantitative variables)

```{r}
plot(gapminder$year, gapminder$gdpPercap)
```

From this scatterplot we can see that GDP per capita is increasing (with the exception of a few outliers in the earlier years)

##### Plotting one quantitative variable

```{r}
ggplot(gapminder, aes(x=lifeExp)) + geom_density()
```

Density plot, which is a smoothed version of the above histogram!

##### Stripchart (one quantitative, one categorical variable)

```{r}
stripchart(gapminder$lifeExp ~ gapminder$continent, vertical=TRUE, method="jitter", jitter=0.1)
```

Made a stripchart instead of a boxplot since I couldn't get the variable names to fit! But as we saw in class, sometimes this data display can be more informative since it includes all data points.

#### Using filter( ), select( ) and %>%

```{r}
myplot <- gapminder %>%
  filter(country=="Canada") %>%
  select(year, pop)
plot(myplot, xlab = "year", ylab = "population", col = "purple", pch = 16)
```

Utilized filter( ), select( ) and %>% to create a plot showing how Canada's population is increasing over time!

```{r}
gapminder %>%
  filter(continent=="Oceania") %>% 
  select(country, year, gdpPercap) %>% 
  ggplot(aes(x=year, y=gdpPercap)) %>% 
  + geom_point(aes(color=country))
```

Fancy graph made by piping into ggplot and using many features!

#### Even more!

```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

The code did not succeed in showing all values for Rwanda and Afghanistan but rather only half the data because no year is repeated, as it should be, showing all years for each region

```{r}
filter(gapminder, country == "Rwanda" | country == "Afghanistan")
```

This is the proper way to do it to obtain all 24 rows of data!

```{r}
knitr::kable(filter(gapminder, country == "Rwanda" | country == "Afghanistan"))
```

Presenting the table in a nicer way using knitr::kable( )

```{r}
gapminder %>%
  filter(year == 1992) %>%
  mutate(logGdp = log10(gdpPercap)) %>% 
  select(continent, lifeExp, logGdp) %>% 
  ggplot(aes(x=logGdp, y=lifeExp)) %>% 
  +geom_point(aes(color=continent))
```

Another, more detailed ggplot graph using an extra dplyr feature.

```{r}
gapminder %>%
  group_by(continent) %>% 
  select(continent, year, pop, gdpPercap) %>% 
  mutate(popgain = pop - first(pop)) %>% 
  filter(continent != "Asia") %>% 
  ggplot(aes(x=gdpPercap, y=popgain)) %>% 
  +geom_point(aes(color=continent))
```

More experimenting with graphs to investigate groupby( ) and to properly use the mutate( ) and first( ) dplyr functions. The graph shows sometimes population increases without gdp per capita (Africa) or vice versa that gdp per capita can grow without population increasing (Europe) or both can increase simultaneously (Americas) and Asia was not included since the scale was a lot larger and it lost the detail of the other four continents.
